---
title: "case_study"
author: "ruby"
date: "2025-10-07"
output: html_document
---

settings/tidyverse 
```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


import dataset
```{r}
library(p8105.datasets)
library(leaflet)

data(nyc_airbnb)
```

## Brainstorm questions 

* investigate price based on neighborhood/borough 
  * does dates affect the price as well?
* does the type of unit affect rating/price?
* what about zip codes?
* what rental has the highest revieews?
* are there any connections between price and other variables 
* prices over time?
  * prices before or after the pandemic?
* how else did the pandemic affect airbnb in nyc?
* does the name of the rental affect the price/rating?
* what does the number of days available have to do with anything else?

airbnb prices by neighbourhood
```{r}
nyc_airbnb |> 
  select(name, neighbourhood_group, price) |> 
  group_by(neighbourhood_group) |> 
  summarize(mean_price = mean(price)) |> 
  ggplot(aes(x=neighbourhood_group, y = mean_price)) +
  geom_col()
```


lets look at prices first
```{r}
nyc_airbnb |> 
  drop_na(price) |> 
  group_by(neighbourhood_group) |> 
  summarize(
    mean_price = mean(price),
    min_price = min(price),
    max_price = max(price)
  )
```

what about room type in a given neighborhood 
```{r}
nyc_airbnb |> 
  drop_na(price) |> 
  group_by(room_type, neighbourhood_group) |> 
  summarize(
    mean_price = mean(price)
  ) |> 
  pivot_wider(
    names_from = room_type, 
    values_from = mean_price
  )
```


counting 
```{r}
nyc_airbnb |> 
  count(neighbourhood_group, room_type)
```

make some plots (remember to look at extreme values...)

```{r}
nyc_airbnb |> 
  ggplot(aes(x=price)) +
  geom_histogram() +
  facet_grid(room_type ~ neighbourhood_group) +
  scale_x_continuous(limits = c(0,1000))
```


look at price of airbnb in manhattan 
```{r}
nyc_airbnb |> 
  filter(neighbourhood_group == "Manhattan", 
         price < 500, 
         room_type == "Entire home/apt") |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) |> 
  ggplot(aes(x = neighbourhood, y = price)) + 
  geom_violin() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

(remember to look at extreme values...)
```{r}
nyc_airbnb |> 
  filter(price > 9000) 
```

```{r}
nyc_airbnb |> 
  filter(calculated_host_listings_count > 30) 
```


keep looking at location 
```{r}
nyc_airbnb |> 
  filter(neighbourhood_group == "Manhattan", 
         price < 1000, 
         room_type == "Entire home/apt") |> 
  ggplot(aes(x= lat, y = long, color = price)) +
  geom_point()
  
```

can we make a better map using `leaflet` package 
```{r}
pal <- colorNumeric(
  palette = "viridis",
  domain = nyc_airbnb$review_scores_location)

nyc_airbnb |> 
  filter(neighbourhood_group == "Manhattan", 
         price < 1000, 
         room_type == "Entire home/apt") |> 
  leaflet() |> 
  addProviderTiles(providers$CartoDB.Positron) |> 
  addCircleMarkers(
    ~lat, ~long, 
    color = ~pal(review_scores_location),
    radius = 2)
```

